---
title: "Class 8: Mini Project"
author: "Idara Ibekwe: A16865157"
format: pdf
---

## About

In today's lab we will work with fine needle aspiration (FNA) of a breast mass from the University of Wisconsin

## Data Import

```{r}
fna.data= "WisconsinCancer(1).csv"
wisc.df <- read.csv("WisconsinCancer (1).csv", row.names=1)
head(wisc.df)
```
>Q1. How many observations/patients/individuals/samples are in this dataset?
569 observations


>Q2. How many Observations have a malignant diagnosis

```{r}
sum(wisc.df$diagnosis == "M")
```
> 212



>Q3. How many variable/features in the data are suffixed with mean

```{r}
ncol(wisc.df)
```

```{r}
colnames(wisc.df)
```

```{r}
inds <- grep("_mean", colnames(wisc.df))
length(inds)
```

```{r}
grep("_mean", colnames(wisc.df), value=T)
```



##Initial Analysis

Before analysis I want to take out the expert diagnoses column (a.k.a the answer) from our dataset. 

```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
head(diagnosis)
```

```{r}
wisc.data <- wisc.df[,-1]
```


##CLustering

We can try kmeans() clustering first

```{r}
km <- kmeans(wisc.data, centers=2)
km$cluster
```
```{r}
table(km$cluster)
```
Cross-table

```{r}
table(km$cluster, diagnosis)
```

Let's try `hclust()` the key input required for `hclust()` is a distance matrix as produced by the `dist()` function. 

```{r}
hc <- hclust(dist(wisc.data))
```

I can make a tree like figure

```{r}
plot(hc)
```

## PCA

Do we need to scale the data?

We can look at the sd of each column (original variable)

```{r}
round(apply(wisc.data, 2, sd))
```

Yes we need to scale. We will run `prcomp()` with `scale=True()`.

```{r}
wisc.pr <- prcomp(wisc.data,scale=TRUE)
summary(wisc.pr)
```

Generate our main PCA plot (score plot, PC1 vs PC2 plot)

```{r}
library(ggplot2)

res <- as.data.frame(wisc.pr$x)
```


```{r}
ggplot(res) +
  aes(PC1, PC2, col=diagnosis) +
  geom_point()

```


## Combining methods

Clustering on PCA results

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust

```{r}
d <- dist(wisc.pr$x[,1:3])
hc <- hclust(d, method= "ward.D2")
plot(hc)
```

To get my clustering result/membership factor I need to "cut" the tree with the `cutree()` function.

```{r}
grps <- cutree(hc, k=2)
```

> Q. How many patients in each group?

```{r}
table(grps)
```


```{r}
plot(res$PC1, res$PC2, col=grps)
```


## Prediction

We can use our PCA result (model) to do predictions, that is take new unseen data and project it onto our new PC variables

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(res$PC1, res$PC2, col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2],labels=c(1,2), col="white")
```

# Summary

Principle component analysis (PCA) is a super useful method for analyzing large data sets. It works by finding new variables (PCs) that capture the most variance from the original variables in your data set. 



